D-Hammer: Efficient Equational Reasoning for Labelled Dirac Notation

### 2025-01-31 01intro_preliminary.tex
update section 3

The next sections introduce the formal language, labelled Dirac notation and the systematical way to reduce labels and prove equalities. 
For this purpose, we use the following motivating example as the main role in our explanations.
Instead of GHZ states, it is a similar but simpler example on Bell states:
\begin{example}
    \label{ex: motivating}
    Let \( q \) and \( r \) represent two quantum systems in the Hilbert space \( \mathcal{H}_T \). Let \( M \) be a quantum operation acting on \( \mathcal{H}_T \), and let \( \ket{\Phi} = \sum_{i \in T} \ket{i} \otimes \ket{i} \) be the maximally entangled state. Then, it holds that
    \[
    M_q \ket{\Phi}_{(q, r)} = M_r^T \ket{\Phi}_{(q, r)}.
    \]
\end{example}
Following the introduction on labels, we can consider the global system $(q,r)$ and transform the equation above into the plain Dirac notation:
\begin{align}
    (M \otimes I) \ket{\Phi} = (I \otimes M^T) \ket{\Phi}.
    \label{eq: motivating plain}
\end{align}
Together with corresponding explanations for this example, we show how an automated system is built and used to solve equalities likewise.

>>>>

In the following,we use the motivating example as the main role in our demonstrations.
Instead of GHZ states, it is a similar but simpler example on Bell states:
\begin{example}
    \label{ex: motivating}
    Let \( q \) and \( r \) represent two quantum systems in the Hilbert space \( \mathcal{H}_T \). Let \( M \) be a quantum operation acting on \( \mathcal{H}_T \), and let \( \ket{\Phi} = \sum_{i \in T} \ket{i} \otimes \ket{i} \) be the maximally entangled state. Then, it holds that
    \[
    M_q \ket{\Phi}_{(q, r)} = M_r^T \ket{\Phi}_{(q, r)}.
    \]
\end{example}
Following the explanation on labels, we can consider the global system $(q,r)$ and transform the equation above into the plain Dirac notation:
\begin{align}
    (M \otimes I) \ket{\Phi} = (I \otimes M^T) \ket{\Phi}.
    \label{eq: motivating plain}
\end{align}
The next sections introduce the formal language, labelled Dirac notation and the systematical way to reduce labels.
We show how an automated system is built and used to solve equalities likewise.

### 2025-01-31 02language_design.tex
update section 3

>>>>

\Cref{fig: normalization demo} shows the normalization outline for $(M \otimes \mathbf{1}_\mathcal{O}(T)) \cdot (\text{phi}\ T)$.

\begin{figure}
    \scriptsize
    \begin{align*}
        & (\eqnmarkbox[blue]{mvar}{M} \otimes \eqnmarkbox[red]{identity}{\mathbf{1}_\mathcal{O}(T)})(\eqnmarkbox[red]{phi-definition}{\textrm{phi}\ T})
        \Rightarrow
        (\sum_{k,l\in T} \bra{k} M \ket{l}.\ket{k}\bra{l}
        \eqnmarkbox[red]{sumpush1}{\otimes \sum_{i \in T}} \ket{i}\bra{i})(\eqnmarkbox[red]{sumpush2}{\sum_{j\in T}}\ket{j, j}) 
        \\
        & \Rightarrow
        \sum_{i,j,k,l\in T} \bra{k} M \ket{l} . \ket{k, i}\eqnmarkbox[red]{braket-dot}{\bra{l, i} \cdot \ket{j,j}}
        \\
        & \Rightarrow
        \eqnmarkbox[red]{delta-elim-sum}{\sum_{i,j,k,l\in T}} (\eqnmarkbox[red]{delta-elim-delta}{\delta_{l,j} \times \delta_{i,j}} \times \bra{k} M \ket{l}) . \ket{k, i}
        \Rightarrow
        \eqnmarkbox[brown]{swapsum}{\sum_{j, k \in T}} \bra{k} M \ket{j} . \ket{k, j}
        \\[1em]
        & \Rightarrow
        \sum_{k, j \in T} \bra{k} M \ket{j} . \ket{k, j}
        \ \eqnmarkbox[brown]{deBruijn}{\xRightarrow{\text{de Bruijn}}}\ 
        \sum_T\sum_T\bra{\$1} M \ket{\$0} . \ket{\$1, \$0}
    \end{align*}
    \annotate[yshift=1em]{right}{mvar}{\scriptsize{variable expansion}}
    \annotate[yshift=0.5em, xshift=1cm]{right}{identity}{\scriptsize{identity operator}}
    \annotate[yshift=0em]{below,left}{phi-definition}{\scriptsize{definition}}
    \annotatetwo[yshift=0em]{below}{sumpush1}{sumpush2}{\scriptsize{sum lifting}}
    \annotate[yshift=0em]{below}{braket-dot}{\scriptsize{inner product}}
    \annotatetwo[yshift=0em,xshift=2cm]{below}{delta-elim-sum}{delta-elim-delta}{\scriptsize{$\delta$-elimination}}
    \annotate[yshift=0em]{below}{swapsum}{\scriptsize{sum swapping}}
    \caption{A nomralization outline for the left-hand side of~\Cref{eq: motivating plain}. Matched subterms are marked with colors. Blue marking represents variable expansion, red marking represents rule applications, and brown marking represents normalization step 4-6.}
    \label{fig: normalization demo}
\end{figure}

% \begin{gather*}
%     \eqnmarkbox[blue]{node1}{GEBV} = \sum_{i}^{n} \eqnmarkbox[red]{node2}{X_{i}} b_{i} \\
%     \eqnmarkbox[green]{node3}{ssssss}
% \end{gather*}

%     \annotate[yshift=1em]{left}{node1}{Genomic Estimated Breeding Value}
%     \annotate[yshift=-1em]{below}{node2}{Genotype at locus i}
%     \annotatetwo[yshift=2em]{above}{node2}{node3}{var 1}
